RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-01-20 versions floris/REIJ_article_FB2.tex
first commit

One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}. This is generally referred to as goal modeling. Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE. Several goal modeling languages have been developed in the last two decades. The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{URN}, which is part of an ITU-T standard, the User Requirements Notation (URN).

A goal model is usually the result of a discussion process between a group of stakeholders and a requirements engineer. For goal models are constructed in a short amount of time, involving stakeholders with a similar background, it is not often necessary to record all of the details of the discussion process that led to the final goal model. %SG1: This sentence doesn't make sense.. Do you mean for the small systems? The next sentence seemed to follow that logic..
However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops. In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.

@@

\begin{enumerate}
\item
It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}. Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model. If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions. As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
\item
Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage. If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
\item
Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost. For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''. Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost. This leads to a poor understanding of the problem and solution domain. In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}. 
\item
It is not possible to reason about changing arguments and their effect on the goal model, and vice versa. A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model. Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments. This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
\end{enumerate}

@@

\section{Background: Goal-oriented Requirements Language and Practical Argumentation}
\label{sect:background}

@@

The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs). %such as performance, cost,  security, and usability. 
URN allows software and requirements engineers to identify  requirements for a proposed or an evolving system and to review such requirements for correctness and completeness. URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements. In this article, we restrict ourselves to GRL.

@@

A variety of links connect the elements in a GRL model. AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element. A contribution link has a qualitative contribution type or a quantitative contribution. Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.

GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$. Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.

\subsubsection{Motivation for using GRL}
\label{sect:background:grl:motivation}

GRL has the existing tools that allow us to achieve requirements four and five of a formal argumentation theory as described in the introduction: (4) the adapted theory must produce repeatable results that can be formalized and implemented. (5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools. GRL is an international standard with well defined syntax and semantics. This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying issue 4. Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation. Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.  According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provide support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details). Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models. This traceability enables direct impact analysis of goal changes on a design and vice versa. This traceability feature seems to be a good fit with the current research: We aim to add traceability links between intentional elements and their underlying arguments. %SG8: what do you mean by current research?

\subsection{Argument Scheme for Practical Reasoning (PRAS)} %SG10: This section is good but make the examples clearer. It has too much concepts but it makes me a bit lost.
\label{sect:background:pras}

Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g. \cite{raz1978,walton1990}) and Artificial Intelligence \cite{bratman1987,atkinson2007}. One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}. The idea is that an instantiation of such a scheme gives a presumptive argument in favour of, for example, taking an action. This argument can then be tested by posing critical questions about, for example, if the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.

A formal approach to argumentative about actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS), as follows. 
\begin{itemize}[leftmargin=22pt]
\item[PRAS]
In the circumstances $R$,\\
agent $Ag$ should perform action $A$,\\
to achieve new circumstances $S$,\\
which will realize goal $G$,\\
which will promote some value $V$.
\end{itemize}

This scheme distinguishes between the state of affairs brought about by an action, the goal (i.e., the desired features in that state of affairs), and the value (i.e., the motivation why those features are desirable). Take, for instance, the example from \cite{vanZee-etal:er2016}: the company (actor) does not have a standard corporate data model (current circumstances), and therefore decides to define such a model (action), leading to a new data model (new circumstances), which realizes the goal to adhere to a corporate data model, which promotes lower costs (value).

In practical reasoning, conclusions which are at one point acceptable can later be rejected because of new information. For example, if we find out that `adhere to corporate data model' is not a realistic goal, we can argue that we should not define such a new data model. Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized. An unfavourable\footnote{Unfavourably can be either 'no', or 'yes', depending on what the question is.} answer to a critical question then identifies a possible counterargument. One critical question is `Does the goal realize the value stated?'. For our example, we can answer this question favourably with a `yes', as one way to lower costs is to adhere to a company's corporate data model \cite{vanZee-etal:er2016}. Another critical question is `Can the desired goal be realized?'. If the answer to this is `no' - for example, there is evidence that `adhere to corporate data model' is not a realistic goal, then we would say the question has been answered unfavourably, and we have a counterargument: `evidence shows that there is not much attention for adherence to a corporate data model, so it can probably not be realized'. Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal or an alternative goal that promotes the same value. For example, in \cite{vanZee-etal:er2016} it is argued that the action `use data model of package applications' realizes the goal `use canonical data model', which also promotes lower costs. This gives us a counterargument that also follows PRAS.

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an emph{argumentation framework} $(Args, Attacks)$, as introduced by Dung~\cite{Dung1995}, where $Args$ is the set of arguments and $Attack$ is an attack relation between arguments. Figure \ref{fig:pras:example1} shows that an argumentation framework that includes the arguments from the above example can be rendered as a graph. The two alternative PRAS instantiations are A1 and A3. These arguments mutually attack each other, as `adhere to corporate data model' (A1) is an alternative to `adhere to canonical data model' (A3) and vice versa. Argument A2 attacks A1, as it questions the validity of the goal `adhere to corporate data model'.

@@

Given an argumentation framework, the acceptability of arguments can be determined according to the semantics of Dung~\cite{Dung1995}. The idea is that we are justified in accepting an argument if it can be successfully defended against attacking counterarguments. In other words, an argument $A \in Args$ is \emph{defeated} by an attacker $B \in Args$ -- where $(A, B) \in Attacks$ -- unless $B$ itself can be defeated. Take, for example, the argumentation framework in Figure~\ref{fig:pras:example}. Argument A2 is undefeated or justified, because it has no attackers. This makes A1 defeated, because one of its attackers, A2, is undefeated. A3 is then also justified, because its only attacker, A1, is defeated by A2. Thus, the set of justified (undefeated) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.

attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two. However, A2 also attacks A1

 where $Args$ is the set of arguments and $Attack$ is an attack relation between arguments. Dung proposes that an argument $A_1\in Args$ is \emph{defeated} by an attacker $A_2\in Args$, unless $A_2$ itself can be defeated. Although this is an appropriate definition of reason about beliefs, it is not for reasoning about action since we are, to some extent, free to choose what we attempt to bring about. Therefore, we may choose to reject an attacker, even if it cannot be defeated, if we believe that the purpose motivating the attacked argument is more important. For instance, suppose an argument $A_2$ stating that a particular resource is expensive attacks an argument $A_1$ for using the resource. However, we may still choose to use it (i.e., accept argument $A_1$), if that would promote a value that we judge more importantly than expense. In order to accommodate this, \emph{valued-based argumentation frameworks} (VAF)~\cite{bench2003persuasion} are used. The idea is that we are given an \emph{argumentation framework}, a set of values $V$, a function $val$ mapping each argument to a value $v\in V$, and a set of \emph{audiences} $a$ to which the arguments are addressed. Each audience is then represented by a strict partial ordering $>_a$ on the values. Then, for a given audience $>_a$, $A_2$ defeats $A_1$ iff $A_2$ attacks $A_1$ and $val(A_2)\not>_a val(A_1)$. \floris{you do not give formal definitions here}

It is then possible to determine which arguments in $Args$ are acceptable for an audience by determining the preferred extension. This is the maximal (under set inclusion) subset $S$ of $Args$ such that no two arguments in $S$ defeat each other, and all arguments in $S$ are acceptable with respect to $S$, i.e., for any argument $A_1$ in $S$, if $A_1$ is defeated by an argument $A_2$ that is not in $S$, then there exists an argument in $S$ that defeats $A_2$. The preferred extension represents the maximal consistent set of acceptable arguments with respect to the argumentation framework and a given audience or value ordering.

\subsubsection{PRAS Medical Treatment Example} %SG11: Isn't it better to use an example more relevant to what we do?
\label{sect:background:pras:example}

A concrete example of using PRAS comes from the medical domain, where Atkinson and colleagues use practical reasoning to formalize a system for reasoning about the treatment of a patient~\cite{atkinson2006argumentation}. The system in this example consists of a group of experts in different areas, such as \emph{treatment knowledge} and \emph{patient knowledge}.\footnote{The original application uses a single expert using different knowledge bases, but we changed this to a group of experts, each with a different expertise, in order to make the comparison with goal modeling more apparent.} There are five critical questions relevant to the application:
\begin{enumerate}
\item[CQ1] Are there alternative ways of realizing the same effects?
\item[CQ2] Are there alternative ways of realizing the same goal?
\item[CQ3] Are the assumptions on which the argument is based true?
\item[CQ4] Does performing the action have side effects which demotes some other value?
\item[CQ5] Will the action have the effects described?
\end{enumerate}

The experts can construct an argumentation framework for a medical treatment by instantiating PAS and posing critical questions CQ1-CQ5. Consider for instance a patient whose health is threatened by blood clotting. The experts start with the null option \--- do nothing (EA0). Then, suppose that the treatment knowledge experts suggests blood clotting can be prevented by reducing platelet adhesion, which can, assuming aspirin is not contraindicated, be achieved by prescribing aspirin. This information is used to instantiate PAS:\\

\noindent
\textbf{EA1.} \emph{Assuming no contradictions, we should prescribe aspirin, which will reduce platelet adhesion, preventing blood clotting, and so is an efficacious course of action.}\\

\noindent
This argument is then subjected to the critical questions. CQ1 asks for alternative solutions to reduce platelet adhesion. The expert may reply by stating chlopidogrel also reduces platelet adhesion. Asking CQ2 forces the expert to seek further solutions for preventing blood clotting, which results in an alternative action, namely to administer streptokinase, which has the same goal of preventing blood clotting. These are formed into two arguments, EA2 and EA3:\\

\noindent
\textbf{EA2.} \emph{Assuming no contradictions, we should prescribe chlopidogrel, which will reduce platelet adhesion, preventing blood clotting, and so is an efficacious course of action.}\\\\
\textbf{EA3}. \emph{Assuming no contradictions, we should prescribe streptokinase, which will increase blood clot dispersal, preventing blood clotting, and so is an efficacious course of action.}\\

\noindent
These three arguments all mutually attack one another, resulting in the argumentation framework shown in figure~\ref{fig:pras:example1}.

Each of the arguments are in mutual conflict, so it is not clear which action to pursue. This depends on the preferences of the experts. For this, suppose the experts vote and agree the effect `reduce platelet adhesion' is a preferred means by which the goal can be realized. Moreover, they prefer aspirin to chlopidogrel because it is cheaper.

These preferences are represented in the argumentation framework by removing attacks of unfavorable actions, so EA1 is no longer attacked, and EA3 no longer attacks EA2. In this case, EA1 forms the preferred extension, and its action is currently the best candidate. However, there are still remaining critical questions that can be asked. For instance, EA1 assumes that aspirin was not contraindicated, and CQ3 instructs us to test this assumption. Suppose that the patient knowledge expert notes the patient has a history of gastritis, and infers that aspirin is contraindicated because its acidity may result in gastric ulceration. The expert forms this into an argument motivated by the value of safety:\\

\noindent
\textbf{EA4.} \emph{Where there is a history of gastritis and no acid reducing therapy, we should not prescribe aspirin, which would cause excess acidity, which would risk ulceration, and so is unsafe.}\\

\begin{figure}[ht]
\centering
\includegraphics[scale=0.4]{img/pras-example2}
\caption{Final argumentation framework showing all critiques from~\cite{atkinson2006argumentation}}
\label{fig:pras:example2}
\end{figure}

\floris{please explain EA5, 6 and 7 as well}
In a similar manner the experts construct three more arguments that attack other arguments in various ways, resulting in the argumentation framework shown in figure~\ref{fig:pras:example2}. In order to determine which arguments are accepted, the preferred extension is computed by first adding the arguments with no attackers, which are EA5, EA6, and EA7. EA7 defeats EA4, which means that EA1 can be included, since its only attacker is defeated. EA1 therefore defeats EA2 and EA4, and also excludes EA0. Thus, in the end, the winning argument is EA1, which prescribes aspirin.

\subsubsection{Motivation for using PRAS}
\label{sect:background:pras:motivation}

In the introduction, we discussed six requirements for a practical reasoning and argumentation framework to be used in our research. The first requirement is: (1) it must have prior successful use in the field of software development. The example of the previous subsection is just one of many where PRAS has been successfully used by computer scientists to formalize arguments in a computer system. \floris{has it been implemented? and does that mean that it has prior successful use in softw dev?} The usefulness of the argument scheme approach for the analysis of practical reasoning situations has been shown in various other areas including e-democracy~\cite{Atkinson2006}, law~\cite{bench2005persuasion} and reasoning about the morally correct course of action~\cite{bench2006automating} \floris{Safety critical actions, explaining character motives in stories, planning}. The second requirement from the introduction is: (2) it should allow to formally reason about the consistency of arguments. The formal semantics in terms of value-based argumentation frameworks allows for the computation of accepted arguments. This semantics is extensively studied in the literature, and various different semantics exists, which lead to different flavors of acceptability of arguments.~\cite{} The third requirement from the introduction is: (3) it must be adaptable for use in goal modeling. In the next section we adapt PRAS in order to do so, but already at first glance the concepts that are used in PRAS and in goal modeling are very similar. In GRL, ``tasks'' can be performed by actors, while in argument scheme PAS these are called ``actions''. Both PAS and GRL have the concept of a ``goal''. Finally, the GRL notion of a ``softgoal'' seems semantically identical to PAS's ``value'', namely that it is a underlying motivation for which goals are desirable. We therefore decided that PRAS does indeed meet our first three requirements of a formal argumentation theory to be used in our research.

%SG12: Read up to the section below.
\section{Argument Schemes for Goal Modeling (GMAS)}
\label{sect:gmas}

PRAS is not directly applicable to goal modeling in general, and GRL in particular. There are differences in the concepts used, and the relationships between these concepts. Some elements appear in PRAS that do not appear in GRL, some elements appear in GRL but not in PRAS, and some concepts appear in both formalisms but with different names. Moreover, GRL allows more relations between elements than the ones appearing in PAS. In this section, we first analyze these differences in more detail. After this, we introduce and motivate Argument Schemes for Goal Modeling (GMAS) and associated critical questions, which are an extension to PRAS. While PRAS contains a single argument scheme PAS, GMAS instead consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model. With each argument scheme we associate critical questions that can be used to attack arguments or provide alternatives. This enables focused discussions about elements and relationships of a goal model. In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.

\subsection{Differences between GRL and PRAS}

>>>>

One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}. This is generally referred to as goal modeling. Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE. Several goal modeling languages have been developed in the last two decades. The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.

A goal model is usually the result of a discussion process between a group of stakeholders and a requirements engineer. For small-sized systems that their goal models are usually constructed in a short amount of time, involving stakeholders with a similar background, it is not often necessary to record all of the details of the discussion process that led to the final goal model. %SG1: I rephrased this since the sentence didn't make any sense logically and grammatically. 
However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops. In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.

@@

\begin{enumerate}
\item
It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}. Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model. If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions. As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
\item
Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage. If these users have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
\item
Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost. For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''. Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost. This leads to a poor understanding of the problem and solution domain. In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}. 
\item
It is not possible to reason about changing arguments and their effect on the goal model, and vice versa. A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model. Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments. This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
\end{enumerate}

@@

\section{Background: Argument Schemes and Goal-oriented Requirements Language}
\label{sect:background}

@@

The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs). %such as performance, cost,  security, and usability. 
URN allows software and requirements engineers to identify  requirements for a proposed or an evolving system and to review such requirements for correctness and completeness. URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements. In this article, we focus on GRL for capturing and documenting argumentation.

@@

A variety of links connect the intentional elements in a GRL model. AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one intentional element on another one. A contribution link has a qualitative contribution type or a quantitative contribution values. Dependency links (\includegraphics[scale=1]{img/dependency}) model the relationships between actors or intentional elements.

GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$. Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution. %SGNew1: I don't think we need this paragraph at all.

\floris{include an example here about the corporate data model etc, which is also used in section \ref{sect:background:pras}. This will make it easier for people to understand the PRAS reasoning}

\subsubsection{Motivation for using GRL}
\label{sect:background:grl:motivation} %SGNew3: I will rewrite this later.

GRL has the existing tools %SGNew2:? what do you mean existing tools? has some means? potential? 
that allow us to achieve requirements four and five of a formal argumentation theory as described in the introduction: (4) the adapted theory must produce repeatable results that can be formalized and implemented. (5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools. GRL is an international standard with well defined syntax and semantics. This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying criterion 4. Moreover, GRL has an open source tool called jUCMNav~\cite{jUCMNav}, which simplifies an implementation. Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL. According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provide support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details). Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models. This traceability enables direct impact analysis of goal changes on a design and vice versa. This traceability feature seems to be a good fit with the current research: We aim to add traceability links between intentional elements and their underlying arguments. %SG8: what do you mean by current research?

\subsection{Argument Scheme for Practical Reasoning (PRAS)} 
\label{sect:background:pras}

Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g. \cite{raz1978, walton1990}) and Artificial Intelligence \cite{bratman1987, atkinson2007}. One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}. The idea is that an instantiation of such a scheme gives a presumptive argument in favor of, for example, taking an action. This argument can then be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.

A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS). PRAS follows the following basic argument structure. 
\begin{equation}
\label{eq:eq1}
  \begin{aligned}
 \qquad&\text{I have goal } G&\\
&\text{Doing action }A \text{ will realize goal }G&\\
&\text{Which will promote value }V&\\
&\text{\emph{Therefore} I should do action }A.&
  \end{aligned}
\end{equation}

So, for example, we can say that XXXX. The argument scheme can be slightly amended to capture subgoals (\emph{i.e.}, realizing goals $G_1, \ldots, G_n$ will allow me to realize goal $G_i$).

Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information. Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements. Some examples of critical questions are as follows.

\begin{enumerate}
\item Will the action bring about the desired goal?
\item Does the goal promote the value stated?
\item Are there alternative ways of realizing the same goal?
\item Are there alternative ways of promoting the same value?
\item Does doing the action have a side effect which demotes the value?
\item Does doing the action have a side effect which demotes some other value?
\item Does doing the action promote some other value?
\item Does doing the action preclude some other action which would promote some other value?
\item Is the action possible?
\item Can the desired goal be realized?
\item Is the value indeed a legitimate value?
\end{enumerate}

An unfavorable\footnote{Unfavorable can be either 'no', or 'yes', depending on what the question is.} answer to a critical question then identifies a possible counterargument. Take, for example, question 2. In the case of our example, we can answer this question favourably with a `yes', as a common way to adhere to a company's corporate data model is obviously to define such a data model. However, take critical question 11: if this is answered negatively -- for example, there is evidence that `adhere to corporate data model' was hardly used in project evaluations over the last four years \cite{vanZee-etal:er2016} -- then we have a counterargument: `evidence shows that there is not much attention for adherence to a corporate data model, so it can probably not be realized'. Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question X) or an alternative goal that promotes the same value (question Z). For example, in \cite{vanZee-etal:er2016} it is argued that the action `use data model of package applications' realizes the goal `use canonical data model', which also promotes lower costs. This essentially gives us a counterargument to the original argument -- to define a corporate data model -- that also follows PRAS.

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}. In this section, we will briefly discuss the intuitions behind these semantics.}. Figure \ref{fig:pras:example1} shows that an argumentation framework that includes the arguments from the above example can be rendered as a graph. The two alternative PRAS instantiations are A1 and A3. These arguments mutually attack each other, as `adhere to corporate data model' (A1) is an alternative to `adhere to canonical data model' (A3) and vice versa. Argument A2 attacks A1, as it questions the validity of the goal `adhere to corporate data model'.

@@

Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics. The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated. Take, for example, the argumentation framework in Figure~\ref{fig:pras:example}. Argument A2 is undefeated because it has no attackers. This makes A1 defeated, because one of its attackers, A2, is undefeated. A3 is then also undefeated, since its only attacker, A1, is defeated by A2. Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.

In some cases, it is more difficult to determine whether or not an argument is defeated. Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two. It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{Amgoud}. Take, for example, A1 and A3. If we say that we prefer the goal ``XXX'' (A1) over the goal ``YYY'' (A3), we remove the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, left). This makes A1 the only undefeated argument, whereas A3 is now defeated. It is also possible to give explicit arguments for preferences \cite{Modgil2009}. These arguments are then essentially attacks on attacks. For example, say we prefer A1 over A3 because ``XXX''. This can be rendered as a separate argument that attacks the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A1, A4$\}$ the undefeated justified set of arguments.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.6]{img/Fig2}
\caption{Preferences between arguments}
\label{fig:pras:example2}
\end{figure}

\subsubsection{Practical Reasoning and Goal Modeling}
\label{sect:background:pras:motivation}

Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung95} and its various extensions \cite{Amgoud, Modgil2009}. The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{Atkinson2006}, law~\cite{???}, planning \cite{MedellinXXX} and choosing between safety critical actions \cite{}. The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{}).

PRAS  and GRL have some obvious similarities in that PRAS defines elements as actions, goals, values and GRL includes tasks, goals, softgoals. In previous work, we have already discussed how to translate practical reasoning arguments to GRL diagrams~\cite{}. For example, ... A similar approach, albeit without the explicit focus on practical reasoning, was taken by Jureta et al.~\cite{}, in which it translates similar structured arguments to goal diagrams. One of the problems of these approaches is that ...

The concept that appears in PRAS and not in GRL is \emph{circumstance}. In PRAS, an action $a$ is performed in current circumstances $R$, and by performing this action new circumstances $S$ are obtained. From this new circumstance $S$, another action may be performed again, and so on. In the AI literature, this type of reasoning usually referred to as ``AI planning''~\cite{weld1999recent}: from some initial state, find a sequence of actions such that a goal state is reached if these actions are executed. This planning approach is taken by a number of contributions using PRAS to create joint plans for a group of actors, often using a transition system as the underlying semantics~\cite{medellin2014}. However, in the early RE phase of an IS, one generally does not consider temporal planning. Decisions about the temporal order in which tasks should be executed are generally postponed until a later stage. In other words, goal modeling is not concerned with making plans, scenarios, or business processes for the information system. In URN (see section~\ref{sect:background:grl}) the distinction between goal modeling and business process modeling is clearly visible in the distinction between two separate modeling languages for the two activities. First, goal models are created in GRL and then, the use case maps (similar to business processes) are created in UCM. Therefore, when extending PRAS for GRL, we do not consider the notion of \emph{circumstances}.

The concept that appears in GRL but not in PRAS is \emph{resource}. A resource may be used by part of the system, or an actor, in order to perform a task or to reach a goal. Therefore, PRAS has to be extended with the notion of a resource in order for it to be applicable to GRL. %SGNew: to be applicable? don't understand the sentence.

Finally, there are two concepts which appear in both models under different names. Where PRAS uses the notions ``action'' and ``value'', GRL uses the concepts ``task'' and respectively ``softgoal''. However, these two concepts are semantically identical.

With respect to relationships between concepts, the argument scheme PRAS from Section~\ref{sect:background:pras} contains the following relationships:\\
